Surrogate Regret Bounds for Polyhedral Losses

### 2021-05-28 rates-neurips-21/rates.tex

\begin{abstract}
  Surrogate risk minimization is an ubiquitous paradigm in supervised machine learning, wherein a target problem is solved by minimizing a surrogate loss on a dataset.
  Surrogate regret bounds are a common tool to prove generalization rates for surrogate risk minimization.
  While surrogate regret bounds have been developed for certain classes of loss functions, such as proper losses, general results are relatively sparse.
  We provide two general results.
  The first gives a linear surrogate regret bound for any polyhedral (piecewise-linear convex) surrogate, meaning that surrogate generalization rates translate directly to target rates.
  The second shows that for sufficiently non-polyhedral surrogates, the regret bound is a square root, meaning fast surrogate generalization rates translate to slow rates for the target.
  Together, these results suggest polyhedral surrogates are weakly optimal in many cases.
  \raft{need to add keyword and/or mention: ``excess risk bounds'' (more common in the literature actually)}
\end{abstract}

@@

We also give a partial converse: If the surrogate is sufficiently ``non-polyhedral'', i.e. smooth and locally strongly convex, then $\zeta$ cannot shrink faster than a square root.
For these losses, therefore, even fast surrogate generalization rates would likely translate to slow rates for the target.
%This result could be viewed as a partial converse to the first, providing evidence that some element of piecewise-linearity is necessary for optimal surrogate regret bounds.
\begin{theorem}[Lower bound for ``non-polyhedral'']
  \label{thm:main-lower}
  Let $L$ be a locally strongly convex surrogate with locally Lipschitz gradient.
  If $L$ satisfies Equation \ref{eq:surrogate-regret-bound-informal} for a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
%  If $\zeta$ is a transfer function in a surrogate regret bound for $L$ to a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
\end{theorem}
% Thus, polyhedral losses are not only order-optimal in terms of regret transfer, but a significant element of piecewise-linearity seems \emph{necesseary} for optimal transfer.
While it may seem intuitive that $\zeta=\sqrt{\cdot}$ for strongly convex losses, it is well-known that losses like exponential loss for binary classification also have $\zeta=\sqrt{\cdot}$, even though they are neither strongly convex nor strongly smooth~\cite{bartlett2006convexity}.
Theorem~\ref{thm:main-lower} offers an explanation for this phenomenon, and together with Theorem~\ref{thm:main-upper} clarifies an intriguing trend in the literature: $\zeta$ is typically either linear or square-root.

@@

\raft{I didn't get much out of this paragraph (apart from the first sentence), whereas I thought the next one was really useful.  I think I'll combine with a shortened version of the first, like ``Theorem 1 in \S, Theorem 2 in \S.  We make Theorem 1 constructive in \S...'' \bo{Sounds good}}
The proof of Theorem \ref{thm:main-upper}, presented in \S~\ref{sec:upper}, relies on techniques from \emph{property elicitation} and the structure of optimal prediction sets for polyhedral losses.
In particular, a polyhedral loss has some finite set of reports which always contains a minimizer~\cite{finocchiaro2019embedding}.\raft{FYI, you were using our embedding dimension paper; I fixed throughout \bo{Thanks; just realized this as well.}}
% This is related to the embeddings framework of \citet{finocchiaro2020embedding}.
Theorem \ref{thm:main-lower}, presented in \S~\ref{sec:lower}, uses curvature of the expected loss to construct a counterexample where convergence of target regret is quadratically slower than that of surrogate regret.

Our proof of the upper bound, Theorem \ref{thm:main-upper}, is nonconstructive.
We make it constructive in Section \ref{sec:constant}.
Specifically, a polyhedral loss's regret transfer function can be bounded by $\zeta(t) \leq \alpha \cdot t$ where the constant is of the form $\alpha = \alpha_{\ell} \cdot \alpha_L \cdot \alpha_{\psi}$.
Each component of $\alpha$ is constructed as a function of its respective object $\ell$, $L$, and $\psi$.
A corollary of the derivation is that a polyhedral loss $L$ and link $\psi$ are consistent for $\ell$ if and only if they satisfy the weak indirect elicitation criterion and $\psi$ is \emph{$\epsilon$-separated}, a condition introduced in \citet{finocchiaro2019embedding}.

@@

The following result is our main lower bound.
\begin{theorem}[Theorem \ref{thm:main-lower} restated.] \label{thm:main-lower-details}
  Suppose the surrogate loss $L$ and link $\psi$ satisfy a regret transfer of $\zeta$ for a target loss $\ell$.
  If $L$, $\psi$, and $\ell$ satisfy Assumption \ref{assumption:lower}, then there exists $c > 0$ such that, for arbitrarily small $\epsilon$, $\zeta(\epsilon) \geq c \sqrt{\epsilon}$.
\end{theorem}
By ``arbitrarily small'', we mean that for any $\epsilon_0 > 0$, we can find $\epsilon$ with $0 < \epsilon < \epsilon_0$ such that $\zeta(\epsilon) \geq c \sqrt{\epsilon}$.
Thus, there is no neighborhood of zero where $\zeta(\epsilon)$ shrinks faster than on the order of $\sqrt{\epsilon}$.

The key idea of the proof is to fix a boundary report $u_0$ and consider a sequence of distributions $p_{\lambda} \to p_0$ as $\lambda \to 0$.
Target regret $R_{\ell}(\psi(u_0),p_{\lambda})$ shrinks linearly in $\lambda$, but surrogate regret $R_L(u_0,p_{\lambda})$ shrinks quadratically.
This holds roughly because expected surrogate loss is strongly convex in some neighborhood of $u_0$ and is strongly smooth elsewhere.\footnote{We note that a tempting approach instead is to fix a distribution $p_0$ and consider a sequence of predictions. But, we know from Lemma \ref{lemma:fixed-p} that this will fail: for any consistent surrogate and any fixed $p_0$, there is a linear regret transfer if we restrict to $p_0$.}

>>>>

\begin{abstract}
  Surrogate risk minimization is an ubiquitous paradigm in supervised machine learning, wherein a target problem is solved by minimizing a surrogate loss on a dataset.
  Surrogate regret bounds are a common tool to prove generalization or excess risk rates for surrogate risk minimization.
  While surrogate regret bounds have been developed for certain classes of loss functions, such as proper losses, general results are relatively sparse.
  We provide two general results.
  The first gives a linear surrogate regret bound for any polyhedral (piecewise-linear convex) surrogate, meaning that surrogate generalization rates translate directly to target rates.
  The second shows that for sufficiently non-polyhedral surrogates, the regret bound is a square root, meaning fast surrogate generalization rates translate to slow rates for the target.
  Together, these results suggest polyhedral surrogates are weakly optimal in many cases.
  \raft{need to add keyword and/or mention: ``excess risk bounds'' (more common in the literature actually) \bo{Did above}}
\end{abstract}

@@

We also give a partial converse: If the surrogate is sufficiently ``non-polyhedral'', i.e. smooth and locally strongly convex, then $\zeta$ cannot shrink faster than a square root.
For these losses, therefore, even fast surrogate generalization rates would likely translate to slow rates for the target.
%This result could be viewed as a partial converse to the first, providing evidence that some element of piecewise-linearity is necessary for optimal surrogate regret bounds.
\begin{theorem}[Lower bound for ``non-polyhedral'']
  \label{thm:main-lower}
  Let $L$ be a locally strongly convex surrogate with locally Lipschitz gradient.
  If $L$ satisfies Equation \ref{eq:surrogate-regret-bound-informal} for a target loss $\ell$, then there exists $c>0$ such that, for all small enough $\epsilon \geq 0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
%  If $\zeta$ is a transfer function in a surrogate regret bound for $L$ to a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
\end{theorem}
% Thus, polyhedral losses are not only order-optimal in terms of regret transfer, but a significant element of piecewise-linearity seems \emph{necesseary} for optimal transfer.
While it may seem intuitive that $\zeta=\sqrt{\cdot}$ for strongly convex losses, it is well-known that losses like exponential loss for binary classification also have $\zeta=\sqrt{\cdot}$, even though they are neither strongly convex nor strongly smooth~\cite{bartlett2006convexity}.
Theorem~\ref{thm:main-lower} offers an explanation for this phenomenon, and together with Theorem~\ref{thm:main-upper} clarifies an intriguing trend in the literature: $\zeta$ is typically either linear or square-root.

@@

\botodo{Commented out a whole paragraph. I don't think we need an outline of the paper.}
%\raft{I didn't get much out of this paragraph (apart from the first sentence), whereas I thought the next one was really useful.  I think I'll combine with a shortened version of the first, like ``Theorem 1 in \S, Theorem 2 in \S.  We make Theorem 1 constructive in \S...'' \bo{Sounds good}}
%The proof of Theorem \ref{thm:main-upper}, presented in \S~\ref{sec:upper}, relies on techniques from \emph{property elicitation} and the structure of optimal prediction sets for polyhedral losses.
%In particular, a polyhedral loss has some finite set of reports which always contains a minimizer~\cite{finocchiaro2019embedding}.\raft{FYI, you were using our embedding dimension paper; I fixed throughout \bo{Thanks; just realized this as well.}}
%% This is related to the embeddings framework of \citet{finocchiaro2020embedding}.
%Theorem \ref{thm:main-lower}, presented in \S~\ref{sec:lower}, uses curvature of the expected loss to construct a counterexample where convergence of target regret is quadratically slower than that of surrogate regret.

Our proof of the upper bound, Theorem \ref{thm:main-upper}, is nonconstructive.
After presenting the main results (Sections \ref{sec:upper} and \ref{sec:lower}), we make it constructive in Section \ref{sec:constant}.
Specifically, a polyhedral loss's regret transfer function can be bounded by $\zeta(t) \leq \alpha \cdot t$ where the constant is of the form $\alpha = \alpha_{\ell} \cdot \alpha_L \cdot \alpha_{\psi}$.
Each component of $\alpha$ is constructed as a function of its respective object $\ell$, $L$, and $\psi$.
A corollary of the derivation is that a polyhedral loss $L$ and link $\psi$ are consistent for $\ell$ if and only if they satisfy the weak indirect elicitation criterion and $\psi$ is \emph{$\epsilon$-separated}, a condition introduced in \citet{finocchiaro2019embedding}.

@@

The following result is our main lower bound.
\begin{theorem}[Theorem \ref{thm:main-lower} restated.] \label{thm:main-lower-details}
  Suppose the surrogate loss $L$ and link $\psi$ satisfy a regret transfer of $\zeta$ for a target loss $\ell$.
  If $L$, $\psi$, and $\ell$ satisfy Assumption \ref{assumption:lower}, then there exists $c > 0$ such that, for some $\epsilon^* > 0$, for all $0 \leq \epsilon < \epsilon^*$, $\zeta(\epsilon) \geq c \sqrt{\epsilon}$.
\end{theorem}
The key idea of the proof is to fix a boundary report $u_0$ and consider a sequence of distributions $p_{\lambda} \to p_0$ as $\lambda \to 0$.
Target regret $R_{\ell}(\psi(u_0),p_{\lambda})$ shrinks linearly in $\lambda$, but surrogate regret $R_L(u_0,p_{\lambda})$ shrinks quadratically.
This holds roughly because expected surrogate loss is strongly convex in some neighborhood of $u_0$ and is strongly smooth elsewhere.\footnote{We note that a tempting approach instead is to fix a distribution $p_0$ and consider a sequence of predictions. But, we know from Lemma \ref{lemma:fixed-p} that this will fail: for any consistent surrogate and any fixed $p_0$, there is a linear regret transfer if we restrict to $p_0$.}
